Embracing the Laws of Physics: \\ Three Reversible Models of Computation

### 2018-07-04 survey/survey.tex
partial work on OpSem section.

\documentclass{article}
\usepackage[utf8x]{inputenc}
\usepackage{agda}
\usepackage[nocenter]{qtree}
\usepackage{amssymb,amsthm,amsmath}
\usepackage{fullpage}
\usepackage{url}
\usepackage{comment}
\usepackage{tikz}
\usepackage{tikz-cd}
\usepackage{listings}
\usepackage{multicol}
\usepackage{amsmath}
\usepackage{stmaryrd}
\usepackage{proof}
\usetikzlibrary{quotes,fit,positioning}
\tikzset{>=latex}

@@

\jc{Maybe interpret the next level as simply term rewriting?
We already know the start and end, but if we also have a whole
sequence of `trans', then we can reconstruct the whole thing
from that. In other words, the example below could be
reconstructed from the start, end and the program itself,
erasing all of the lhs annotations, which are merely there for
humans to track.}

>>>>

\documentclass{article}
\usepackage[utf8x]{inputenc}
\usepackage{agda}
\usepackage[nocenter]{qtree}
\usepackage{amssymb,amsthm,amsmath}
\usepackage{fullpage}
\usepackage{url}
\usepackage{comment}
\usepackage{tikz}
\usepackage{tikz-cd}
\usepackage{listings}
\usepackage{multicol}
\usepackage{stmaryrd}
\usepackage{proof}
\usetikzlibrary{quotes,fit,positioning}
\tikzset{>=latex}

@@

There are two different interpretations for an operational semantics
for the language of equivalences:
\begin{enumerate}
\item Mimicking closely the one in the previous section, and
thus finding explicit homotopies between the functions induced
by the operational semantics of the level-1 combinators.
\item Treating things more syntactically, and interpreting
the combinators as program transformations.
\end{enumerate}
A previous paper~\cite{Carette2016} explores the first interpretation
in depth. There one can find a definition of ``equivalences of
equivalences'', which as the base of that interpretation.

Here we will focus instead of the syntactic interpretation as
program transformers. This results in a function
\begin{flalign*}
\mathit{eval}_1\,:\,\ \left\{ t_1\, t_2\,:\,U\right\}
  \left\{ c_1\, c_2\,:\, t_1\,\leftrightarrow\,t_2\right\}
  (ce\,\,c_1 \Leftrightarrow\,c_2) \rightarrow (t_1\,\leftrightarrow\,t_2) \\
\end{flalign*}
This function is ``deeply dependent'': given the type of the
rewrite $ce$ to apply, both the input $c_1$ and output $c_2$ are
almost entirely determined!  Let us take for example the
second combinator in Fig.~\ref{figi}:
\[
  {(c_1 \otimes (c_2 \oplus c_3)) \odot \distl \Leftrightarrow \distl \odot ((c_1 \otimes c_2) \oplus (c_1 \otimes c_3))}
\]
\noindent which we can name \AgdaInductiveConstructor{distl⇔l}.
Interpreting this as a rewrite from the program on the left to the
one on the right requires ``pattern matching'' on the left
structure which contains $3$ arbitrary combinators, from which
we can \emph{reconstruct} the program on the right. Rewrites
such as \AgdaInductiveConstructor{distl⇔l} are one-step rewrites,
in the same way that \distl is a constant of the base term language
of $\Pi$.  There is one additional wrinkle. There is naturally
an opposite combinator, which interprets the above from right
to left; let us call it \AgdaInductiveConstructor{distl⇔r}. It
would appear to require \emph{non-linear pattern-matching}
since the right-hand-side contains $c_1$ twice. That is however not
the case! The definition of \AgdaInductiveConstructor{distl⇔r} has
$5$ implicit arguments, $3$ of which are $c_1, c_2$, and $c_3$,
which then completely force the ``shape'' of the overall pattern.
Thus the mere mention of \AgdaInductiveConstructor{distl⇔r} is
enough to resolve the apparent use of a non-linear pattern.
This is why $\mathit{eval}_1$ was called ``deeply dependent''
above: once the name of the combinator is given, the rest
follows.
